### 2023-10-16 appendix_3.Rmd

---
appendix: "appendix_3.Rmd"
---

## Simulated Parital Invariance Results

```{r p-tab11, results = 'asis', eval = T}
partial.invariant <-
  partial_mi(
    saved_model = results.invariant$invariance_models$model.residuals,
    data = df.invariant,
    model = model.overall,
    group = "group",
    group.equal = c("loadings", "intercepts", "residuals"),
    partial_step = "residuals")

apa_table(partial.invariant$fit_table %>% 
  dplyr::select(free.parameter, cfi, rmsea), 
  digits = 3, 
  col.names = c("Estimated Parameter", "CFI", "RSMEA"),
  caption = "Fit Estimates for Partial Invariance Residuals on Invariant Data")
```

```{r p-tab12, results = 'asis', eval = T}
partial.small.load <-
  partial_mi(
    saved_model = results.small.load$invariance_models$model.loadings,
    data = df.small.load,
    model = model.overall,
    group = "group",
    group.equal = c("loadings", "intercepts", "residuals"),
    partial_step = "loadings")

apa_table(partial.small.load$fit_table %>% 
  dplyr::select(free.parameter, cfi, rmsea), 
  digits = 3, 
  col.names = c("Estimated Parameter", "CFI", "RSMEA"),
  caption = "Fit Estimates for Partial Invariance Loadings for Small Loading Data")
```

```{r p-tab13, results = 'asis', eval = T}
partial.med.load <-
  partial_mi(
    saved_model = results.med.load$invariance_models$model.loadings,
    data = df.med.load,
    model = model.overall,
    group = "group",
    group.equal = c("loadings", "intercepts", "residuals"),
    partial_step = "loadings")

apa_table(partial.med.load$fit_table %>% 
  dplyr::select(free.parameter, cfi, rmsea), 
  digits = 3, 
  col.names = c("Estimated Parameter", "CFI", "RSMEA"),
  caption = "Fit Estimates for Partial Invariance Loadings for Medium Loading Data")
```

```{r p-tab14, results = 'asis', eval = T}
partial.large.load <-
  partial_mi(
    saved_model = results.large.load$invariance_models$model.loadings,
    data = df.large.load,
    model = model.overall,
    group = "group",
    group.equal = c("loadings", "intercepts", "residuals"),
    partial_step = "loadings")

apa_table(partial.large.load$fit_table %>% 
  dplyr::select(free.parameter, cfi, rmsea), 
  digits = 3, 
  col.names = c("Estimated Parameter", "CFI", "RSMEA"),
  caption = "Fit Estimates for Partial Invariance Loadings for Large Loading Data")
```

```{r p-tab15, results = 'asis', eval = T}
partial.small.int <-
  partial_mi(
    saved_model = results.small.int$invariance_models$model.intercepts,
    data = df.small.int,
    model = model.overall,
    group = "group",
    group.equal = c("loadings", "intercepts", "residuals"),
    partial_step = "intercepts")

apa_table(partial.small.int$fit_table %>% 
  dplyr::select(free.parameter, cfi, rmsea), 
  digits = 3, 
  col.names = c("Estimated Parameter", "CFI", "RSMEA"),
  caption = "Fit Estimates for Partial Invariance Loadings for Small Intercept Data")
```

```{r p-tab16, results = 'asis', eval = T}
partial.med.int <-
  partial_mi(
    saved_model = results.med.int$invariance_models$model.intercepts,
    data = df.med.int,
    model = model.overall,
    group = "group",
    group.equal = c("loadings", "intercepts", "residuals"),
    partial_step = "intercepts")

apa_table(partial.med.int$fit_table %>% 
  dplyr::select(free.parameter, cfi, rmsea), 
  digits = 3, 
  col.names = c("Estimated Parameter", "CFI", "RSMEA"),
  caption = "Fit Estimates for Partial Invariance Loadings for Medium Intercept Data")
```

```{r p-tab17, results = 'asis', eval = T}
partial.large.int <-
  partial_mi(
    saved_model = results.large.int$invariance_models$model.intercepts,
    data = df.large.int,
    model = model.overall,
    group = "group",
    group.equal = c("loadings", "intercepts", "residuals"),
    partial_step = "intercepts")

apa_table(partial.large.int$fit_table %>% 
  dplyr::select(free.parameter, cfi, rmsea), 
  digits = 3, 
  col.names = c("Estimated Parameter", "CFI", "RSMEA"),
  caption = "Fit Estimates for Partial Invariance Loadings for Large Intercept Data")
```

```{r p-tab18, results = 'asis', eval = T}
partial.small.res <-
  partial_mi(
    saved_model = results.small.res$invariance_models$model.residuals,
    data = df.small.res,
    model = model.overall,
    group = "group",
    group.equal = c("loadings", "intercepts", "residuals"),
    partial_step = "residuals")

apa_table(partial.small.res$fit_table %>% 
  dplyr::select(free.parameter, cfi, rmsea), 
  digits = 3, 
  col.names = c("Estimated Parameter", "CFI", "RSMEA"),
  caption = "Fit Estimates for Partial Invariance Loadings for Small Residual Data")
```

```{r p-tab19, results = 'asis', eval = T}
partial.med.res <-
  partial_mi(
    saved_model = results.med.res$invariance_models$model.residuals,
    data = df.med.res,
    model = model.overall,
    group = "group",
    group.equal = c("loadings", "intercepts", "residuals"),
    partial_step = "residuals")

apa_table(partial.med.res$fit_table %>% 
  dplyr::select(free.parameter, cfi, rmsea), 
  digits = 3, 
  col.names = c("Estimated Parameter", "CFI", "RSMEA"),
  caption = "Fit Estimates for Partial Invariance Loadings for Medium Residual Data")
```

```{r p-tab20, results = 'asis', eval = T}
partial.large.res <-
  partial_mi(
    saved_model = results.large.res$invariance_models$model.residuals,
    data = df.large.res,
    model = model.overall,
    group = "group",
    group.equal = c("loadings", "intercepts", "residuals"),
    partial_step = "residuals")

apa_table(partial.large.res$fit_table %>% 
  dplyr::select(free.parameter, cfi, rmsea), 
  digits = 3, 
  col.names = c("Estimated Parameter", "CFI", "RSMEA"),
  caption = "Fit Estimates for Partial Invariance Loadings for Large Residual Data")
```

\newpage