### 2023-10-16 appendix_4.Rmd

---
appendix: "appendix_4.Rmd"
---
## Invariance Plots Difference Scores by Condition

```{r inv-plots-condition, include = TRUE}
conditions <- unique(boot_long$type)

for (cond in conditions){
  invariance_plot <- ggplot(boot_long %>% 
                              filter(type == cond), 
                          aes(term, difference, color = type_estimate)) + 
  stat_summary(fun = mean, geom = "point", 
               position = position_dodge(width = 0.9)) + 
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", 
               position = position_dodge(width = 0.9), 
               width = 0.2) + 
  theme_classic() + 
  xlab("Estimated Item") + 
  ylab("Difference Score Between Groups") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), legend.position = "bottom") + 
  facet_wrap(~index_difference, labeller = as_labeller(label_graph)) + 
  scale_color_discrete(name = "Type of Estimate") + 
    coord_flip() + 
    ggtitle(paste0(cond, " Graph Results"))

print(invariance_plot)
}
```