### 2023-10-16 appendix_5.Rmd

---
appendix: "appendix_5.Rmd"
---

## Invariance Plots Effect Sizes by Condition

```{r inv-effect-condition, include = TRUE}

for (cond in conditions){
  effect_invariance_plot <- 
  ggplot(boot_summary_long %>% 
           filter(type == cond), aes(term, d, color = type_estimate)) + 
  theme_classic() + 
  facet_wrap(~invariant, labeller = as_labeller(label_graph)) +
  scale_color_discrete(name = "Type of Estimate") + 
  geom_point(aes(size = n)) + 
  geom_errorbar(aes(ymin = d_low, ymax = d_high), width = 0.2) + 
  xlab("Estimated Item") + 
  ylab("Group Effect Size") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), 
        legend.position = "bottom") + 
  geom_hline(yintercept = 0) + 
  coord_flip() + 
    ggtitle(paste0(cond, " Graph Results"))

  print(effect_invariance_plot)
}
```