### 2024-07-29 appendix_7.Rmd

---
appendix: "appendix_7.Rmd"
---

## Replication Test

```{r setup}
knitr::opts_chunk$set(echo = TRUE, include = TRUE,
                      message = FALSE, warning = FALSE, size = "small")
```

```{r libraries-appendix7, message = FALSE}
set.seed(890343)
library(rio)
library(dplyr)
library(visualizemi)
library(papaja)
library(tidyr)
library(dmacs)
library(ggplot2)
library(scales)
library(introdataviz)
library(lavaan)
```

## Data

```{r}
# our data
load("manu_data/RS14.Rdata")

DF <- DF

# take only interesting columns and give a label
DF <- DF %>% 
  # make it comparable by only testing students
  filter(DF$sample == "Student") %>% 
  mutate(group = "Aiena") %>% 
  dplyr::select(RS1:RS14, group)

# Chen et al.
DF.pone <- import("manu_data/data1-RS14.dta")
# take only interesting columns and give a label
DF.pone <- DF.pone %>% 
  dplyr::select(starts_with("rs")) %>% 
  mutate(group = "Chen")
# match column style
colnames(DF.pone)[1:14] <- toupper(colnames(DF.pone)[1:14])

# create one dataset
DF.combo <- bind_rows(DF, DF.pone)

DF.combo %>% 
  group_by(group) %>% 
  summarize(sample = n())
```

## MGCFA

```{r}
# build the one-factor model 
model.rs <- "RS =~ RS1+RS2+RS3+RS4+RS5+RS6+RS7+RS8+RS9+RS10+RS11+RS12+RS13+RS14"

# run the multi-group CFA
results.rs <- mgcfa(
  model = model.rs,
  data = DF.combo, 
  group = "group", 
  group.equal = c("loadings", "intercepts", "residuals")
  )

# examine the results
results.rs$model_fit %>% 
        dplyr::select(model, AIC, BIC, cfi, tli, rmsea, srmr)
```

Overall, the one-factor model fits the data well. Each group also shows adequate model fit. If we use $\Delta$CFI <= .01, we find that the loadings would be considered invariant across the English and Chinese samples. The intercepts were not invariant.

## Partial Invariance

```{r}
partial.rs <-
  partial_mi(
    saved_model = results.rs$invariance_models$model.intercepts,
    data = DF.combo,
    model = model.rs,
    group = "group",
    # be sure to do only up to the step you are interested in
    group.equal = c("loadings", "intercepts"),
    partial_step = "intercepts")

partial.rs$fit_table %>% 
  dplyr::select(free.parameter, cfi) %>% 
  arrange(-cfi)
```

Examining partial invariance reveals several potential candidates for partial invariance. In this next section, we relaxed group constraints until we achieved partial invariance (i.e., $\Delta$CFI <= .01). We will need to find our CFI as at least `r results.rs$model_fit %>% filter(model == "Configural") %>% pull(cfi) %>% apa_num() %>% as.numeric() - .01`. More than half the items are necessary to achieve "partial" invariance (which really implies no invariance is likely possible).

```{r}
# run the partially invariant model with group.partial
partial.rs.1 <- mgcfa(model = model.rs, 
                  data = DF.combo, 
                  group = "group", 
                  group.equal = c("loadings", "intercepts"),
                  group.partial = c("RS2 ~1", "RS11 ~1",
                                    "RS4 ~1", "RS12 ~1",
                                    "RS14 ~1", "RS1 ~1",
                                    "RS10 ~1", "RS3 ~ 1"),
                  meanstructure = TRUE)

# examine the fit indices 
partial.rs.1$model_fit %>% 
  filter(model == "intercepts") %>% 
  dplyr::select(AIC, BIC, cfi, tli, rmsea, srmr)

# examine the intercepts 
partial.rs.1$model_coef %>% 
  filter(model == "intercepts") %>% 
  dplyr::select(term, group, estimate) %>% 
  pivot_wider(id_cols = c("term"), 
              names_from = "group", 
              values_from = "estimate") %>% 
  dplyr::rename(English = `1`, 
                Chinese = `2`) %>% 
  filter(term %in% c("RS1 ~1 ", "RS12 ~1 ",
                     "RS7 ~1 ", "RS4 ~1 ",
                     "RS5 ~1 ", "RS3 ~1 "))

# effect size model 
round(lavaan_dmacs(partial.rs.1$invariance_models$model.intercepts, "Chen")$DMACS, 2)
```

## Visualize Invariance

```{r}
for (i in c("RS2", "RS11",
            "RS4", "RS12",
            "RS14", "RS1",
            "RS10", "RS3")){
  cat(paste0("Plot for ", i, "\n"))
  print(plot_mi(
    # output from model_coef
    data_coef = partial.rs.1$model_coef, 
    # which model do you want to plot
    model_step = "intercepts", 
    # name of observed item
    item_name = i, 
    # latent variable limits to graph
    x_limits = c(-1,1), 
    # Y min and max in data 
    y_limits = c(min(DF.combo[ , i]), max(DF.combo[ , i])),
    # what ci do you want
    conf.level = .95, 
    # what model results do you want 
    model_results = partial.rs.1$invariance_models$model.intercepts,
    # which latent variable do you want 
    lv_name = "RS" 
)$complete)

  ggsave(paste0("figures/rs14-chinese-", i, ".png"), 
         dpi = 300, 
         width = 8, 
         units = "in")

  cat(paste0("-------\n"))

}

```

## Bootstrap Model

```{r eval = F}
boot.model.rs <- 
  bootstrap_model(
    saved_configural = results.rs$model_configural,
    data = DF.combo,
    model = model.rs, 
    group = "group", 
    nboot = 1000, 
    invariance_index = "cfi",
    invariance_rule = .01, 
    group.equal = c("loadings", "intercepts")
)

# export(boot.model.rs, "manu_data/boot.model.rs.appendix7.csv")

boot.model.rs
```

```{r echo = F}
boot.model.rs <- import("manu_data/boot.model.rs.appendix7.csv")
boot.model.rs %>% 
  dplyr::select(model, non_invariant, random_non_invariant, 
         h_nmi, h_nmi_p)
```

In this case, we do not see loadings print out. That implies that all models in both real data and randomized data are invariant because the function only calculates information for non-invariance. We see that the intercepts are unlikely to ever replicate across Chinese and English samples. This result is not surprising given the large number of relaxed parameters required to achieve partial invariance.

## Bootstrap Partial Invariance

```{r eval = F, message = F}
boot.partial.rs <- 
  bootstrap_partial(
    saved_model = results.rs$invariance_models$model.intercepts,
    data = DF.combo,
    model = model.rs, 
    group = "group", 
    nboot = 1000, 
    invariance_index = "cfi",
    invariance_rule = .01, 
    invariance_compare = fitmeasures(results.rs$invariance_models$model.loadings, "cfi"), 
    partial_step = "intercepts",
    group.equal = c("loadings", "intercepts")
)

saveRDS(boot.partial.rs, "manu_data/boot.partial.rs.appendix7.Rdata")
```

```{r echo = F}
load("manu_data/boot.partial.rs.appendix7.Rdata")
boot.partial.rs <- list(
  "boot_DF" = boot_DF,
  "boot_summary" = boot_summary,
  "density_plot" = density_plot, 
  "invariance_plot" = invariance_plot, 
  "boot_effects" = boot_effects, 
  "effect_invariance_plot" = effect_invariance_plot
)
```

### Each Parameter on the Overall Model Invariance

```{r}
boot.partial.rs$boot_effects %>% 
  dplyr::select(term, non_invariant, random_non_invariant, 
         h_nmi, h_nmi_p)
```

In this output, we see that all the bootstrapped runs of the real data are non-invariant, even when each parameter is relaxed individually. A few runs of the random data are *non*-invariant (meaning most are actually invariant when randomized). This indicates that no one parameter is likely the reason for non-invariance, as they all show large non-replication effects. If we use the `boot_summary`, we can see the effect size for each parameter when the two intercepts are compared to each other (as the chart above shows the overall model invariance effect).

### Each Parameter's Standardized Difference Score

```{r}
boot.partial.rs$boot_summary %>% 
  dplyr::select(term, invariant, n_boot, d_boot)

boot.partial.rs$boot_summary %>% 
  dplyr::select(term, invariant, n_random, d_random)
```

With the bootstrap summary, we see that no invariant intercept runs are found (not all columns shown to save space). Note that the `RS~1` does not calculate *d* because the *sd* is zero (thus, this would cause an error as *sd* is part of the denominator). The random runs also do not show a *d* score because the effect size is only calculated when at least 10% of the runs in `n_boot` or `n_random` are found. Since we only have a few in that column, no effect size is calculated. The differences in the intercepts for the real data (`boot` columns) appear to be medium to large, showing *d* scores from 0.3 to 0.6. Therefore, we might expect that the English and Chinese samples have different average endorsement levels of the RS14.

```{r}
boot.partial.rs$density_plot

ggsave(paste0("figures/rs14-chinese-density_plot.png"), 
       dpi = 300, 
       width = 8, 
       units = "in")
```

The examination of the density plot shows how group 1 (Aiena English) tends to show lower average scores than group 2 (Chen Chinese) for most but not all items.

We can view the mean difference or standardized mean difference by using:

```{r}
boot.partial.rs$invariance_plot

ggsave(paste0("figures/rs14-chinese-invariance_plot.png"), 
       dpi = 300, 
       width = 8, 
       units = "in")

boot.partial.rs$effect_invariance_plot

ggsave(paste0("figures/rs14-chinese-effect_invariance_plot.png"), 
       dpi = 300, 
       width = 8, 
       units = "in")
```

When effects are non-invariant in the randomized data, the mean difference is still fairly small, but we see large mean differences in intercepts when the bootstrapped data is non-invariant. In the effect size graph, we can see that this effect is medium to large for all the parameters.